The language of generalization: Probability, vagueness, and interaction

### 2018-05-08 paper/appendix-cueValidity.Rmd
refactor cue validity appendix to load rather than compute results

```{r echo = F}
library(langcog)
library(tidyverse)
library(ggthemes)
library(jsonlite)
theme_set(theme_few())
project.path <- "../"
data.path <- "data/generics/endorsement/"
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
```

@@

```{r directQuestion, echo = F}
d.cv.dq <- read.csv(paste(project.path, data.path, "cue-validity-1-trials.csv", sep = ""))

@@

d.cv.targets <- left_join(
  d.cv.dq %>%
    mutate(property = gsub("&quotechar", "'", property), 
         item = paste(category, property)),
  target.items) %>%
  filter(!is.na(type)) %>%
  mutate(type = factor(type, levels = c("diagnostic", "undiagnostic", "false")))

d.cv.summary <- d.cv.targets %>%
  group_by(item, type) %>%
  multi_boot_standard(col = "response")

# ggplot(d.cv.summary, aes ( x = item, y = mean, ymin = ci_lower, ymax = ci_upper))+
#   geom_bar(stat = 'identity', position = position_dodge())+
#   geom_errorbar(position = position_dodge())+
#   facet_wrap(~type, scales = 'free')+
#   scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
#   theme(axis.text.x = element_text(angle = 90))+
#   xlab("")+
#   ylab("Probability rating of cue validity")+
#   ggtitle("Direct question paradigm")

@@

```{r freeProduction, echo = F}
d.cv.fp <- read.csv(paste(project.path, 
                          data.path, "cue-validity-2-freeProduction-trials-n100.csv", sep = ""))

mosquito.mispellings <- c("mosqu", "mesqu", "misqu", "mosiq")

d.cv.fp <- left_join(
  d.cv.fp %>%
 # rename(property = category) %>%
  mutate(response = tolower(response),
         response = gsub(" ", "", response),
         response = ifelse(substr(response, 1, 5) %in% mosquito.mispellings, "mosquito", response),
         response = ifelse(response == "deertick", "tick", response),
         response = ifelse(substrRight(response, 1) == "s",
                           substr(response, 1, nchar(response)-1), response),
         property = gsub("&quotechar", "'", property),
         item = paste(response, property)),
    target.items)

d.cv.fp.summary <- left_join(
  target.items,
  d.cv.fp %>%
    rename(category = response) %>%
    group_by(property) %>%
    mutate(n = n()) %>%
    filter(!(is.na(type))) %>%
    group_by(category, property, item) %>%
    summarize(mentions = n(),
              trials = mean(n), # mean(n) == n, because it's just the number of subjects
              prop = mentions / trials)
  ) %>%
  mutate(prop = ifelse(is.na(prop), 0.01, prop)) %>%
  mutate(type = factor(type, levels = c("diagnostic", "undiagnostic", "false")))

d.cv.bootstrapped <- data.frame()
resample_n <- length(levels(factor(d.cv.fp$workerid)))
for (i in 1:100){

  d.cv.bsample <-left_join(
    target.items,
    d.cv.fp %>% 
    select(workerid, property, response) %>% 
    spread(property, response) %>% 
    sample_n(resample_n, replace = TRUE) %>%
    gather(property, category, -workerid) %>% 
    group_by(property) %>%
    mutate(n = n(),
           item = paste(category, property))
    ) %>%
    group_by(category, property, item, type) %>%
    summarize(mentions = n(),
              trials = mean(n), # mean(n) == n, because it's just the number of subjects
              prop = mentions / trials,
              prop = ifelse(is.na(prop), 0.01, prop),
              iteration = i)

  d.cv.bootstrapped <- bind_rows(d.cv.bootstrapped, d.cv.bsample)
}

empiricalLower = function(dist){
  xi <- quantile(dist, 0.025)
  return (xi[["2.5%"]])
}
empiricalUpper = function(dist){
  xi <- quantile(dist, 0.975)
  return (xi[["97.5%"]])
}
empiricalMean = function(dist){
  xi <- quantile(dist, 0.5)
  return (xi[["50%"]])
}

d.cv.fp.summary <- d.cv.bootstrapped %>%
      group_by(item, type) %>%
      summarize(mean = empiricalMean(prop),
                ci_lower = empiricalLower(prop),
                ci_upper = empiricalUpper(prop)) %>%
  mutate(
    ci_lower = ifelse(is.na(mean), 0.01, ci_lower),
    ci_upper = ifelse(is.na(mean), 0.01, ci_upper),
    mean = ifelse(is.na(mean), 0.01, mean)
  )

d.cv.bothtasks <- bind_rows(
  d.cv.summary %>% mutate(task = 'direct question'),
  d.cv.fp.summary %>% mutate(task = 'free production')
)

```{r}
cv.bothquestions.bars <- d.cv.bothtasks %>%
  ungroup() %>%
  mutate(type = factor(type, levels = c( "false", "undiagnostic","diagnostic"))) %>%
  ggplot(., aes ( x = item, y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_bar(stat = 'identity', position = position_dodge(0.5),
           width = 0.5, color = 'black', fill = 'white')+
  geom_errorbar(position = position_dodge(0.5), width = 0.3)+
  facet_grid(task~type, scales = 'free')+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  geom_hline(data = theoretical.cv, aes(yintercept = cv), lty = 2)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        strip.text.y = element_text(angle = 0))+
  xlab("")+
  ylab("Cue validity")+
  guides(color = F)
```

```{r dq.histogram}
dq.histograms <- ggplot(d.cv.targets %>%
         filter(item %in% undiagnostic.props$item), aes( x = response ))+
  geom_histogram() + 
  facet_wrap(~item, nrow = 2) + 
  scale_x_continuous(breaks = c(0, 1))+
  scale_y_continuous(breaks = c(0, 10))+
  xlab("Direct question response")+
  ylab("Number of responses")
```

```{r}
diagnostic.props <- data.frame(
  item = c("kangaroo has a pouch", "lion has a mane", 
           "mosquito carries malaria", "tick carries Lyme disease"),
  type = "diagnostic"
)

undiagnostic.props <- data.frame(
  item = c("leopard is juvenile", "lion is male", "robin is female", "swan is full-grown"),
  type = "undiagnostic"
)

target.items2 <- bind_rows(diagnostic.props, undiagnostic.props, false.props)
```

```{r cv-fp-data-analyzeAll, cache = T}
d.target.items <- fromJSON(paste(project.path, "data/generics/endorsement/", "originalStims.json", sep = "")) %>%
  mutate(
    property = gsub("'", "", property),
    Property = gsub("'", "", Property),
    sentence = paste(Category, Property),
    item = paste(category, property))

d.cv.bootstrapped.all <- data.frame()
resample_n <- length(levels(factor(d.cv.fp$workerid)))
for (i in 1:1000){

  d.cv.bsample <- d.cv.fp %>% 
    select(workerid, property, response) %>% 
    spread(property, response) %>% 
    sample_n(resample_n, replace = TRUE) %>%
    gather(property, category, -workerid) %>% 
        group_by(property) %>%
    mutate(n = n(),
           item = paste(category, property)) %>%
    filter(item %in% d.target.items$item) %>%
    group_by(category, property) %>%
    summarize(mentions = n(),
              trials = mean(n), # mean(n) == n, because it's just the number of subjects
              prop = mentions / trials,
              prop = ifelse(is.na(prop), 0.01, prop),
              iteration = i)

  d.cv.bootstrapped.all <- bind_rows(d.cv.bootstrapped.all, d.cv.bsample)
}

d.gen.cv.all.summary <- left_join(
    d.target.items %>% select(-sentence),
    d.cv.bootstrapped.all %>%
      group_by(category, property) %>%
      summarize(cv_mean = empiricalMean(prop),
                cv_ci_lower = empiricalLower(prop),
                cv_ci_upper = empiricalUpper(prop))
) %>%
  mutate(
    cv_ci_lower = ifelse(is.na(cv_mean), 0.01, cv_ci_lower),
    cv_ci_upper = ifelse(is.na(cv_mean), 0.01, cv_ci_upper),
    cv_mean = ifelse(is.na(cv_mean), 0.01, cv_mean)
  )

```{r cuevalidity-model-pk}
wp.pkg.path <- paste(project.path, "models/generics/node_modules/utils/", sep = "")

cueValidityModel1 <- '
var cueValidity = function(){
  var kind = uniformDraw(data.prior);
  observe(Bernoulli({p: kind.prevalence/100}), true)
  return kind.Category == data.kind[0]
}

expectation(Infer({model: cueValidity}))
'

d.prev <- read.csv(paste(project.path, "data/generics/endorsement/",
                       "naturalGenerics-prior-trials-n57.csv", sep = ""))

d.gen.endorse.priors <- read.csv(paste(project.path, 
                                       "data/generics/endorsement/",
                                       "naturalGenerics-prior-trials-n57.csv", sep = ""))
gen.endorse.properties <- levels(d.gen.endorse.priors$Property)

wp.cueValidity <- data.frame()
for (p in gen.endorse.properties){
  categories <- levels(factor(filter(d.gen.cv.all.summary, Property == p)$Category))
  prevPriorData = d.gen.endorse.priors %>% filter(Property == p)

  for (catg in categories){

    dataToPass <- list(
      prior = prevPriorData,
      kind = catg
    )

    wp.rs <- webppl(cueValidityModel1, 
       data_var = "data",
       data = dataToPass)

    wp.cueValidity <- bind_rows(wp.cueValidity, 
                                data.frame(Category = catg,
                                           Property = p,
                                           cv = wp.rs))

  }
  #print(p)
}
# save(wp.cueValidity, file =
#        paste(project.path, model.path,
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
# load(file = 
#        paste(project.path, model.path, 
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
```

```{r}
md.fp <- left_join(left_join(
  d.gen.cv.all.summary,
  wp.cueValidity %>% rename(prev_prior_derived_cv = cv)
  #wp.cueValidity.uniform %>% rename(prev_prior_derived_cv = cv)
), target.items2) %>%
  mutate(type = ifelse(is.na(type), "other", type),
           type = factor(type, levels = c("diagnostic", "undiagnostic", "false", "other"),
                       labels = c("diagnostic", "undiagnostic", "false", "other")))

fill.colors <- RColorBrewer::brewer.pal(3, "Set1")

# ggplot(md.fp, aes (x = prev_prior_derived_cv, y = cv_mean, ymin = cv_ci_lower, ymax = cv_ci_upper,
#                    color = type))+
#   #scale_alpha_manual(values = c(1,1,1,0.5))+
#   geom_errorbar(alpha = 0.4)+
#   scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
#   scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
#   geom_point(data = md.fp %>% filter(type == "other"), size = 2)+
#   geom_point(data = md.fp %>% filter(!(type == "other")), size = 2)+
#   coord_fixed()+
#   #scale_color_solarized()+
#   scale_color_manual(values = c(fill.colors, "grey8"))+
#   xlab("Prevalence Prior Derived Cue Validity")+
#   ylab("Free Production Measured Cue Validity")+
#   theme(legend.title = element_blank())

@@

r2.fp.desid <- compute_r2(md.fp.desid, "cv_mean", "prev_prior_derived_cv")
mse.fp.desid <- compute_mse(md.fp.desid, "cv_mean", "prev_prior_derived_cv")
n.fp.desid <- length(md.fp.desid$item)
```

```{r}
d.cv.dq.all <- d.cv.dq %>%
  mutate(property = gsub("&quotechar", "", property), 
         item = paste(category, property)) %>%
  group_by(category, property) %>%
  multi_boot_standard(col = 'response')

md.dq <-  left_join( left_join(
    left_join(
    d.gen.cv.all.summary %>% select(Category, Property, category, property, item),
    d.cv.dq.all),
    #wp.cueValidity.uniform %>% rename(prev_prior_derived_cv = cv)
    wp.cueValidity %>% rename(prev_prior_derived_cv = cv)
  ), target.items2) %>%
  mutate(type = ifelse(is.na(type), "other", type),
           type = factor(type, levels = c("diagnostic", "undiagnostic", "false", "other"),
                       labels = c("diagnostic", "undiagnostic", "false", "other")))

# ggplot(md.dq, aes (x = prev_prior_derived_cv, y = mean, ymin = ci_lower, ymax = ci_upper,
#                      color = type))+
#   geom_errorbar(alpha = 0.4)+
#   scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
#   scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
#   geom_point(data = md.dq %>% filter(type == "other"), size = 2)+
#   geom_point(data = md.dq %>% filter(!(type == "other")), size = 2)+
#   coord_fixed()+
#   #scale_color_solarized()+
#   scale_color_manual(values = c(fill.colors, "grey8"))+
#   xlab("Prevalence Prior Derived Cue Validity")+
#   ylab("Direct Question Measured Cue Validity")+
#   theme(legend.title = element_blank())
# 
#   
r2.dq <- compute_r2(md.dq, "mean", "prev_prior_derived_cv")
mse.dq <- compute_mse(md.dq, "mean", "prev_prior_derived_cv")
n.dq <- length(md.dq$item)

@@

prevprior.cv.scatters <- bind_rows(
  md.fp %>%
    rename(mean = cv_mean, ci_lower = cv_ci_lower, ci_upper = cv_ci_upper) %>%
    mutate(src = "Free production"),
  md.dq %>%
    mutate(src = "Direct question")
) %>%
  ggplot(., aes (x = prev_prior_derived_cv, y = mean, ymin = ci_lower, ymax = ci_upper,
                     color = type))+
  geom_errorbar(alpha = 0.4)+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  geom_point(data = . %>% filter(type == "other"), size = 2)+
  geom_point(data = . %>% filter(!(type == "other")), size = 2)+
  coord_fixed()+
  #scale_color_solarized()+
  scale_color_manual(values = c(fill.colors, "grey8"))+
  xlab("Prevalence Prior Derived Cue Validity")+
  ylab("Measured Cue Validity")+
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.direction = 'horizontal')+
  facet_wrap(~src)
```

```{r cv-bothquestions-barplots, fig.height = 8, fig.width = 8, fig.cap="Top: Empirically measured cue validity for two different tasks. Items are grouped by whether the property is never present in the category (false), the property is always present in the category and every other category (undiagnostic), or present in the category and absent from most other categories (diagnostic). Dotted lines denote theoretical cue validity representing the desiderata (see text). Error bars denote bootstrapped 95% confidence intervals. Bottom left: Raw empirical distributions for direct question task. Bottom right: Correspondence between measured cue validity and prevalence prior derived cue validity."}
grid.arrange(cv.bothquestions.bars, 
             dq.histograms, 
             prevprior.cv.scatters,
             ncol = 4,
             layout_matrix = cbind(c(1,1,2,2),
                                   c(1,1,2,2),
                                   c(1,1,3,3),
                                   c(1,1,3,3)
                                   ))
```

@@

```{r cvModel-uniformPrior, eval = F}
cueValidityModel2 <- '
var cueValidity = function(){
  var kind = uniformDraw(data.allKinds);
  var responses = _.filter(data.prior, {Category: kind});
  var response = _.isEmpty(responses) ? {prevalence: 0} : uniformDraw(responses);
  observe(Bernoulli({p: response.prevalence/100}), true)
  return kind == data.kind[0]
}

expectation(Infer({model: cueValidity}))
'

allKinds = levels(factor(d.gen.endorse.priors$Category))

wp.cueValidity.uniform <- data.frame()
for (p in gen.endorse.properties){
  categories <- levels(factor(filter(d.gen.cv.all.summary, Property == p)$Category))
  prevPriorData = d.gen.endorse.priors %>% filter(Property == p)

  for (catg in categories){

    dataToPass <- list(
      prior = prevPriorData,
      kind = filter(prevPriorData, Category == catg)[1, "Category"],
      allKinds = allKinds
    )

    wp.rs <- webppl(cueValidityModel2, data_var = "data", data = dataToPass)

    wp.cueValidity.uniform <- bind_rows(wp.cueValidity.uniform, 
                                data.frame(Category = catg,
                                           Property = p,
                                           cv = wp.rs))

  }
  print(p)
}
# save(wp.cueValidity, file = 
#        paste(project.path, model.path, 
#              "results/prevPriorDerivedCueValidity.Rdata", sep = ""))
# dataToPass %>% toJSON(.)
# 
# wp.rs

>>>>

```{r}
fill.colors <- RColorBrewer::brewer.pal(3, "Set1")
```

@@

```{r cv-bothquestions-barplots, fig.height = 8, fig.width = 8, fig.cap="Top: Empirically measured cue validity for two different tasks. Items are grouped by whether the property is never present in the category (false), the property is always present in the category and every other category (undiagnostic), or present in the category and absent from most other categories (diagnostic). Dotted lines denote theoretical cue validity representing the desiderata (see text). Error bars denote bootstrapped 95% confidence intervals. Bottom left: Raw empirical distributions for direct question task. Bottom right: Correspondence between measured cue validity and prevalence prior derived cue validity."}
load("cached_results/appendix_cueValidity_results.RData")
# d.cv.bothtasks, d.cv.targets, undiagnostic.props
# md.fp, md.dq

@@

cv.bothquestions.bars <- d.cv.bothtasks %>%
  ungroup() %>%
  mutate(type = factor(type, levels = c( "false", "undiagnostic","diagnostic"))) %>%
  ggplot(., aes ( x = item, y = mean, ymin = ci_lower, ymax = ci_upper))+
  geom_bar(stat = 'identity', position = position_dodge(0.5),
           width = 0.5, color = 'black', fill = 'white')+
  geom_errorbar(position = position_dodge(0.5), width = 0.3)+
  facet_grid(task~type, scales = 'free')+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  geom_hline(data = theoretical.cv, aes(yintercept = cv), lty = 2)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        strip.text.y = element_text(angle = 0))+
  xlab("")+
  ylab("Cue validity")+
  guides(color = F)

dq.histograms <- ggplot(d.cv.targets %>%
         filter(item %in% undiagnostic.props$item), aes( x = response ))+
  geom_histogram() + 
  facet_wrap(~item, nrow = 2) + 
  scale_x_continuous(breaks = c(0, 1))+
  scale_y_continuous(breaks = c(0, 10))+
  xlab("Direct question response")+
  ylab("Number of responses")

diagnostic.props2 <- data.frame(
  item = c("kangaroo has a pouch", "lion has a mane", 
           "mosquito carries malaria", "tick carries Lyme disease"),
  type = "diagnostic"
)

undiagnostic.props2 <- data.frame(
  item = c("leopard is juvenile", "lion is male", "robin is female", "swan is full-grown"),
  type = "undiagnostic"
)

false.props2 <- data.frame(
  item = c("leopard has wings", "kangaroo has spots", "tiger has a pouch", "shark has a mane"),
  type = "false"
)

target.items2 <- bind_rows(diagnostic.props2, undiagnostic.props2, false.props2)

prevprior.cv.scatters <- bind_rows(
  md.fp %>%
    rename(mean = cv_mean, ci_lower = cv_ci_lower, ci_upper = cv_ci_upper) %>%
    mutate(src = "Free production"),
  md.dq %>%
    mutate(src = "Direct question")
) %>%
  ggplot(., aes (x = prev_prior_derived_cv, y = mean, ymin = ci_lower, ymax = ci_upper,
                     color = type))+
  geom_errorbar(alpha = 0.4)+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  geom_point(data = . %>% filter(type == "other"), size = 2)+
  geom_point(data = . %>% filter(!(type == "other")), size = 2)+
  coord_fixed()+
  #scale_color_solarized()+
  scale_color_manual(values = c(fill.colors, "grey8"))+
  xlab("Prevalence Prior Derived Cue Validity")+
  ylab("Measured Cue Validity")+
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.direction = 'horizontal')+
  facet_wrap(~src)

grid.arrange(cv.bothquestions.bars, 
             dq.histograms, 
             prevprior.cv.scatters,
             ncol = 4,
             layout_matrix = cbind(c(1,1,2,2),
                                   c(1,1,2,2),
                                   c(1,1,3,3),
                                   c(1,1,3,3)
                                   ))

@@

```{r}
# d.cv.bothtasks, d.cv.targets, undiagnostic.props
# md.fp, md.dq

@@

r2.fp.desid <- compute_r2(md.fp.desid, "cv_mean", "prev_prior_derived_cv")
mse.fp.desid <- compute_mse(md.fp.desid, "cv_mean", "prev_prior_derived_cv")
n.fp.desid <- length(md.fp.desid$item)

r2.dq <- compute_r2(md.dq, "mean", "prev_prior_derived_cv")
mse.dq <- compute_mse(md.dq, "mean", "prev_prior_derived_cv")
n.dq <- length(md.dq$item)

@@

@@
